---
title: "shiny tutorial"
author: "James Stanfield"
date: "7/26/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```


# Basics of building a shiny app

Template - blank app
```{r}
library(shiny)
ui <- fluidPage()

server <- function(input, output) {}

shinyApp(ui = ui, server = server)
```


Print "Hello World" inside app
```{r}
library(shiny)
ui <- fluidPage("Hello World")

server <- function(input, output) {}

shinyApp(ui = ui, server = server)
```

Input a number slider
```{r}
library(shiny)
ui <- fluidPage(
  sliderInput(inputId = "num",
              label = "Choose a number",
              value = 25, min = 1, max = 100)
)

server <- function(input, output) {}

shinyApp(ui = ui, server = server)
```

List of some other inputs:
```{r}
actionButton()
submitButton()

checkboxInput()
checkboxGroupInput()

dateInput()
dateRangeInput()

fileInput()

numericInput()

passwordInput()

radioButtons()

selectInput()

sliderInput()

textInput()
```


Input syntax
```{r}
#typeofInput(inputId = "unique name of input", label = "label of input seen by user", specific arguments)

#example:

sliderInput(inputId = "num", label = "Choose a number", value = 25, min = 1, max = 100)
```

Some types of outputs:
```{r}
dataTableOutput()
htmlOutput()
imageOutput()
plotOutput()
tableOutput()
textOutput()
uiOutput()
verbatimTextOutput()
```

syntax for output
```{r}
#typeofOutput(outputId = "unique name of output object")

#ex.

plotOutput(outputId = "hist")
```

Add an output command
```{r}
library(shiny)
ui <- fluidPage(
  sliderInput(inputId = "num",
              label = "Choose a number",
              value = 25, min = 1, max = 100),
  plotOutput("hist") #tells shiny to make space for a plot, but not enough for any display on its own
)

server <- function(input, output) {}

shinyApp(ui = ui, server = server)
```


Rules for using server to make convert inputs to outputs
```{r}
server <- function(input, output) {
  output$hist <- #code                1: save objects to display to output$
  output$hist <- renderPlot({code}) # 2: Build objects to display with render*()
  input$num                         # 3: Use input values with input$
}

#syntax example

server <- function(input, output) {
  ouput$hist <- renderPlot({
    hist(rnorm(input$num))
  })
}
```

List of some render functions:
```{r}
renderDataTable()
renderImage()
renderPlot()
renderPrint()
renderTable()
renderText()
renderUI()
```

Combine input with output
```{r}
library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "num", 
    label = "Choose a number", 
    value = 25, min = 1, max = 100),
  plotOutput("hist")
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(rnorm(input$num))
  })
}

shinyApp(ui = ui, server = server) 
```

share apps with
https://www.shinyapps.io/ #make account
https://shiny.rstudio.com/articles/shinyapps.html #post apps on rstudio

https://rstudio.com/products/shiny/shiny-server/  #make own server for apps


# Reactivity

1st reactive function
```{r}
#render functions

renderDataTable()
renderImage()
renderPlot()
renderPrint()
renderTable()
renderText()
renderUI()
```


```{r}
# 01-two-inputs

library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "num", 
    label = "Choose a number", 
    value = 25, min = 1, max = 100),
  textInput(inputId = "title", 
    label = "Write a title",
    value = "Histogram of Random Normal Values"),
  plotOutput("hist")
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(rnorm(input$num), main = input$title)
  })
}

shinyApp(ui = ui, server = server)
```

```{r}
# 02-two-outputs

library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "num", 
    label = "Choose a number", 
    value = 25, min = 1, max = 100),
  plotOutput("hist"),
  verbatimTextOutput("stats")
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(rnorm(input$num))
  })
  output$stats <- renderPrint({
    summary(rnorm(input$num))
  })
}

shinyApp(ui = ui, server = server)
```


2nd function
```{r}
#reactive() 

data <- reactive( {rnorm(input$num)} )

#will cause object to respond to every reactive value in the code

# 1: call a reactive expression like a function i.e. data()

# 2: Reactive expressions cache their values, returning the most recent value unless it has become invalidated.
```

The power of reactive is you can run a process data once, and then use the saved value(s) for any number of outputs.
So when you have multiple outputs, place as much in reactive() as you can in order to streamline both the coding and computation time.

```{r}
# 03-reactive

library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "num", 
    label = "Choose a number", 
    value = 25, min = 1, max = 100),
  plotOutput("hist"),
  verbatimTextOutput("stats")
)

server <- function(input, output) {
  
  data <- reactive({
    rnorm(input$num)
  })
  
  output$hist <- renderPlot({
    hist(data())
  })
  output$stats <- renderPrint({
    summary(data())
  })
}

shinyApp(ui = ui, server = server)
```


3rd function
```{r}
# isolate()

isolate({ rnorm(input$num) })

# returns the result as a non-reactive value
```

```{r}
# 04-isolate

library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "num", 
    label = "Choose a number", 
    value = 25, min = 1, max = 100),
  textInput(inputId = "title", 
    label = "Write a title",
    value = "Histogram of Random Normal Values"),
  plotOutput("hist")
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(rnorm(input$num), main = isolate(input$title))
  })
}

shinyApp(ui = ui, server = server)
```


4th function
```{r}
#observeEvent(reactive values to respond to, { code block to run whenever observer is invalidated })
                                                  #note: code block treated as if it is isolated

observeEvent(input$clicks, { print(input$clicks) })
```

```{r}
actionButton(inputId = "go", label = "Click Me!")
```

```{r}
# 05-actionButton/observeEvent

library(shiny)

ui <- fluidPage(
  actionButton(inputId = "clicks", 
    label = "Click me")
)

server <- function(input, output) {
  observeEvent(input$clicks, {
    print(as.numeric(input$clicks))
  })
}

shinyApp(ui = ui, server = server)
```

```{r}
# 06-observeEvent

library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "num", 
    label = "Choose a number",
    min = 1, max = 100, value = 25),
  actionButton(inputId = "go", 
    label = "Print Value")
)

server <- function(input, output) {
  
  # observe responds to the print button
  # but not the slider
  observeEvent(input$go, {
    print(as.numeric(input$num))
  })
}

shinyApp(ui = ui, server = server)
```


5th function
```{r}
#observe()

#observe({code block to run whenever observer is invalidated})

observe( { print(input$clicks) } )

#also triggers code to run on server.
#same syntax as render, reactive, and isolate
#will respoond to every reactive value
```


6th function
```{r}
# eventReactive()

# data <- eventReactive(reactive value(s) to respond to, { code used to build object })
                                                          #note: object code treated as isolated

data <- eventReactive(input$go, {rnorm(input$num)})

# Delays reactions
# update buttons
```

```{r}
# 07-eventReactive

library(shiny)

ui <- fluidPage(
  sliderInput(inputId = "num", 
    label = "Choose a number", 
    value = 25, min = 1, max = 100),
  actionButton(inputId = "go", 
    label = "Update"),
  plotOutput("hist")
)

server <- function(input, output) {
  data <- eventReactive(input$go, {
    rnorm(input$num) 
  })
  
  output$hist <- renderPlot({
    hist(data())
  })
}

shinyApp(ui = ui, server = server)
```


7th function
```{r}
# reactiveValues()

# rv <- reactiveValues(optional elements to add to the list)

rv <- reactiveValues(data = rnorm(100))

# Creates list of reactive values to manipulate programmatically
```

```{r}
# 08-reactiveValues

library(shiny)

ui <- fluidPage(
  actionButton(inputId = "norm", label = "Normal"),
  actionButton(inputId = "unif", label = "Uniform"),
  plotOutput("hist")
)

server <- function(input, output) {

  rv <- reactiveValues(data = rnorm(100))

  observeEvent(input$norm, { rv$data <- rnorm(100) })
  observeEvent(input$unif, { rv$data <- runif(100) })

  output$hist <- renderPlot({ 
    hist(rv$data) 
  })
}

shinyApp(ui = ui, server = server)
```

Code outside server function will only be run once per session. Maximize code here

Code inside server function will be run once per user. Bridge between Max and Min.

Code inside a reactive function will be run every time the app is interacted with. Minimize code here


Interesting article about rshiny reactivity:
https://shiny.rstudio.com/articles/understanding-reactivity.html 


# Customize Appearance

Tags
```{r}
# tags$h1()  same as html <h1></h1>
# tags$a()   same as html <a></a>
```

```{r}
library(shiny)
names(tags) #run to get list
```

Be sure to follow tags$h1 with (), or you will just get a function description

```{r}
tags$h1
tags$h1()
```

Tags syntax
```{r}
#listy of names tags$function/tag name(name = arguments, unnamed arguments)
#   Note: named arguments appear as tag attributes, unnamed arguments appear inside the tag

tags$a(href = "www.rstudio.com", "Rstudio")

#html:  <a href = "www.rstudio.com">Rstudio</a>
```

Headers
```{r}
fluidPage(
  tags$h1("First level"),
  tags$h2("Second level"),
  tags$h3("Third level"),
  tags$h4("Fourth level"),
  tags$h5("Fifth level"),
  tags$h6("Sixth level")
)
```

hyperlink
```{r}
fluidPage(
tags$a(href = "http://www.git.com", "Git")
)
```

text
```{r}
fluidPage(
  "This is a Shiny app.",
  "It is also a web page."
)
```

paragraphs
```{r}
fluidPage(
  tags$p("This is a Shiny app."),
  tags$p("It is also a web page.")
)
```

Italic and Bold, etc.
```{r}
fluidPage(
  tags$em("I'm italicized!"),
  tags$strong("I'm Bold!"),
  tags$code("I look like code!")
)
```

Nest style tags
```{r}
fluidPage(
  tags$p("This is a", tags$strong("Shiny"), "app.")
)
```

line break
```{r}
fluidPage(
  "This is a Shiny app.",
  tags$br(),
  "It is also a web page."
)
```

Horizontal rule
```{r}
fluidPage(
  "This is a Shiny app.",
  tags$hr(),
  "It is also a web page."
)
```

Insert images
```{r}
fluidPage(
  tags$img(
    height = 100,
    src = "http://www.rstudio.com/images/Rstudio.2x.png" #src = source of image
  )
)
```

add images from own system
```{r}
#create folder inside the same folder as the app, and call it www and place desired images in that folder

fluidPage(
  tags$img(
    height = 100,
    src = "bigorb.png" #r will pull from the www folder, so only include the name of the image, w/subdirectory if needed
  )
)
```

can use html
```{r}
HTML('HTML code')
```

break app into rows
```{r}
fluidPage(
  fluidRow(),
  fluidRow()
)
```

column
```{r}
column(width = , offset = ) #there are 12 units across to assign
#place inside of rows
```

```{r}
# 02-tags.R

library(shiny)

ui <- fluidPage(
  h1("My Shiny App"),
  p(style = "font-family:Impact",
    "See other apps in the",
    a("Shiny Showcase",
      href = "http://www.rstudio.com/
      products/shiny/shiny-user-showcase/")
  )
)

server <- function(input, output){}

shinyApp(ui = ui, server = server)
```

```{r}
# 03-layout.R

library(shiny)

ui <- fluidPage(
  fluidRow(
   column(3),
   column(5, sliderInput(inputId = "num", 
     label = "Choose a number", 
     value = 25, min = 1, max = 100))
  ),
  fluidRow(
    column(4, offset = 8,
      plotOutput("hist")
    )
  )
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(rnorm(input$num))
  })
}

shinyApp(ui = ui, server = server)
```

```{r}
# 04-well.R

library(shiny)

ui <- fluidPage(
  wellPanel(
    sliderInput(inputId = "num", 
      label = "Choose a number", 
      value = 25, min = 1, max = 100),
    textInput(inputId = "title", 
      label = "Write a title",
      value = "Histogram of Random Normal Values")
  ),
  plotOutput("hist")
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(rnorm(input$num), main = input$title)
  })
}

shinyApp(ui = ui, server = server)
```

```{r}
# 05-tabs.R

library(shiny)

ui <- fluidPage(title = "Random generator",
  tabsetPanel(              
    tabPanel(title = "Normal data",
      plotOutput("norm"),
      actionButton("renorm", "Resample")
    ),
    tabPanel(title = "Uniform data",
      plotOutput("unif"),
      actionButton("reunif", "Resample")
    ),
    tabPanel(title = "Chi Squared data",
      plotOutput("chisq"),
      actionButton("rechisq", "Resample")
    )
  )
)

server <- function(input, output) {
  
  rv <- reactiveValues(
    norm = rnorm(500), 
    unif = runif(500),
    chisq = rchisq(500, 2))
  
  observeEvent(input$renorm, { rv$norm <- rnorm(500) })
  observeEvent(input$reunif, { rv$unif <- runif(500) })
  observeEvent(input$rechisq, { rv$chisq <- rchisq(500, 2) })
  
  output$norm <- renderPlot({
    hist(rv$norm, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard normal distribution")
  })
  output$unif <- renderPlot({
    hist(rv$unif, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard uniform distribution")
  })
  output$chisq <- renderPlot({
    hist(rv$chisq, breaks = 30, col = "grey", border = "white",
       main = "500 random draws from a Chi Square distribution with two degree of freedom")
  })
}

shinyApp(server = server, ui = ui)
#© 2020 GitHub, Inc.
```

```{r}
# 06-navlist.R

library(shiny)

ui <- fluidPage(title = "Random generator",
  navlistPanel(              
    tabPanel(title = "Normal data",
      plotOutput("norm"),
      actionButton("renorm", "Resample")
    ),
    tabPanel(title = "Uniform data",
      plotOutput("unif"),
      actionButton("reunif", "Resample")
    ),
    tabPanel(title = "Chi Squared data",
      plotOutput("chisq"),
      actionButton("rechisq", "Resample")
    )
  )
)

server <- function(input, output) {
  
  rv <- reactiveValues(
    norm = rnorm(500), 
    unif = runif(500),
    chisq = rchisq(500, 2))
  
  observeEvent(input$renorm, { rv$norm <- rnorm(500) })
  observeEvent(input$reunif, { rv$unif <- runif(500) })
  observeEvent(input$rechisq, { rv$chisq <- rchisq(500, 2) })
  
  output$norm <- renderPlot({
    hist(rv$norm, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard normal distribution")
  })
  output$unif <- renderPlot({
    hist(rv$unif, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard uniform distribution")
  })
  output$chisq <- renderPlot({
    hist(rv$chisq, breaks = 30, col = "grey", border = "white",
       main = "500 random draws from a Chi Square distribution with two degree of freedom")
  })
}

shinyApp(server = server, ui = ui)
```

```{r}
# 07-Sidebar.R

library(shiny)

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      sliderInput(inputId = "num", 
        label = "Choose a number", 
        value = 25, min = 1, max = 100),
      textInput(inputId = "title", 
        label = "Write a title",
        value = "Histogram of Random Normal Values")
    ),
    mainPanel(
      plotOutput("hist")
    )
  )
)

server <- function(input, output) {
  output$hist <- renderPlot({
    hist(rnorm(input$num), main = input$title)
  })
}

shinyApp(ui = ui, server = server)
```

```{r}
# 08-navbarPage.R

library(shiny)

ui <- navbarPage(title = "Random generator",
    tabPanel(title = "Normal data",
      plotOutput("norm"),
      actionButton("renorm", "Resample")
    ),
    tabPanel(title = "Uniform data",
      plotOutput("unif"),
      actionButton("reunif", "Resample")
    ),
    tabPanel(title = "Chi Squared data",
      plotOutput("chisq"),
      actionButton("rechisq", "Resample")
    )

)

server <- function(input, output) {
  
  rv <- reactiveValues(
    norm = rnorm(500), 
    unif = runif(500),
    chisq = rchisq(500, 2))
  
  observeEvent(input$renorm, { rv$norm <- rnorm(500) })
  observeEvent(input$reunif, { rv$unif <- runif(500) })
  observeEvent(input$rechisq, { rv$chisq <- rchisq(500, 2) })
  
  output$norm <- renderPlot({
    hist(rv$norm, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard normal distribution")
  })
  output$unif <- renderPlot({
    hist(rv$unif, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard uniform distribution")
  })
  output$chisq <- renderPlot({
    hist(rv$chisq, breaks = 30, col = "grey", border = "white",
       main = "500 random draws from a Chi Square distribution with two degree of freedom")
  })
}

shinyApp(server = server, ui = ui)
```

```{r}
# 09-navbarMenu.R

library(shiny)

ui <- navbarPage(title = "Random generator",
  tabPanel(title = "Normal data",
    plotOutput("norm"),
    actionButton("renorm", "Resample")
  ),
  navbarMenu(title = "Other data",
    tabPanel(title = "Uniform data",
      plotOutput("unif"),
      actionButton("reunif", "Resample")
    ),
    tabPanel(title = "Chi Squared data",
      plotOutput("chisq"),
      actionButton("rechisq", "Resample")
    )
  )
)

server <- function(input, output) {
  
  rv <- reactiveValues(
    norm = rnorm(500), 
    unif = runif(500),
    chisq = rchisq(500, 2))
  
  observeEvent(input$renorm, { rv$norm <- rnorm(500) })
  observeEvent(input$reunif, { rv$unif <- runif(500) })
  observeEvent(input$rechisq, { rv$chisq <- rchisq(500, 2) })
  
  output$norm <- renderPlot({
    hist(rv$norm, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard normal distribution")
  })
  output$unif <- renderPlot({
    hist(rv$unif, breaks = 30, col = "grey", border = "white",
      main = "500 random draws from a standard uniform distribution")
  })
  output$chisq <- renderPlot({
    hist(rv$chisq, breaks = 30, col = "grey", border = "white",
       main = "500 random draws from a Chi Square distribution with two degree of freedom")
  })
}

shinyApp(server = server, ui = ui)
```

```{r}

```






























